---
title: "Parts Simulation Study"
author: "Clay Morrow"
date: "5/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(ggplot2)
```


Simulation Model:

$$x(t) = \beta_1 u_1 + \beta_2 u_2 + R_1 r_1 + (p_1 u_1 + p_2 u_2 + R_2 r_2) t + (g_1 + R_3 r_3) w_1(t) + \varepsilon(t) $$

* **x(t)**: pixel value at time **t**

* **u**: land classes

* **r**: random spatially correlated effect (unmeasured; 2D sinwave): makes other effects random

* **t**: time point {1, 2, ..., **T**}.

* **w**: spatio-temporal predictor variable (tapered AR, fixed nugget and range parameter)

* **beta**, **p**, **g**, **R**: constant throughout sim studies. 

* **e(t)**: ARMA(1,1) error: 

$$\varepsilon_i(t) = \delta_i(t) + \rho_i \varepsilon_i(t - 1)  + \theta_i \varepsilon_i(t-1)$$

  - **rho**: constant, except in sim study 4
  
  - **delta**: spatially correlated (tapered, fixed nugget and range parameter)
  
Simulation Studies:

```{r custom_functions}
# spherical tapering function ----
taper.sphr <- function(d, theta){
  ts = ifelse(d > theta, 0, ((1 - d/theta)^2) * (1 + d/(2 * theta)) )
  return(ts)
}
```

```{r taper_sphr_plot, fig.asp = 2/3, fig.width = 5}
# show taper spherical function with varying thetas
curve(taper.sphr(x, .02), from = 0, to = 0.1, n = 101, col = "black",
      main = "Tapering spherical covariance function",
      ylab = "f(distance, range)", xlab = "distance (% of map)")

curve(taper.sphr(x, .04), from = 0, to = 0.1, n = 101, add = TRUE, col = "blue")
curve(taper.sphr(x, .06), from = 0, to = 0.1, n = 101, add = TRUE, col = "forestgreen")
curve(taper.sphr(x, .08), from = 0, to = 0.1, n = 101, add = TRUE, col = "darkorchid3")

legend(x = .075, y = 1, 
       legend = c("0.02", "0.04", "0.06", "0.08"), 
       col = c("black", "blue", "forestgreen", "darkorchid3"), 
       lty = 1, 
       title = "r")
```



  
```{r}
## Select map dimensions that are divisible by 2, 4, and 8

x <- 100:300
y <- x^2

div2 <- x %% 2 == 0
div4 <- x %% 4 == 0
div8 <- x %% 8 == 0

indx <- which(div2 & div4 & div8)

row.sizes = c(x[indx][which(abs(y[indx] - 10000) == min(abs(y[indx] - 10000)))],
              x[indx][which(abs(y[indx] - 20000) == min(abs(y[indx] - 20000)))],
              x[indx][which(abs(y[indx] - 40000) == min(abs(y[indx] - 40000)))],
              x[indx][which(abs(y[indx] - 80000) == min(abs(y[indx] - 80000)))])

# rm(x, y, div2, div4, div8)

# row.sizes/2
# row.sizes/4
# row.sizes/8
```
  
  
```{r sim_one_setup}
# Map parameters ----

# map.rows <- c(100, 150, 200, 280) # map row combinations
# stopifnot(all((map.rows^2) %% 4 == 0)) # check that the square is divisible by 4



map.rows = 16
map.cols = map.rows
npix = map.rows * map.cols

# simulation paramters ----
spat.range = .05 # effective range
nugget = .10
xsd = 1

# Map calculations ----
## X, Y coordinates
coords = as.matrix(expand.grid(x = as.double(1:map.cols), 
                               y = as.double(1:map.rows)))
## Distance
D.time = system.time(D <- scales::rescale(as.matrix(dist(coords)), to = c(0, 1)))

## Covariance matrix
Sigma <- diag(rep(nugget, npix)) + (1 - nugget) * taper.sphr(D, spat.range)

## Cholesky matrix
# Chol.time = system.time(Cholmat <- base::chol(Sigma))

Chol.time2 = system.time(invchol <- remotePARTS::invert_chol(Sigma, nugget))
```

```{r sim_one_landclass, fig.asp = 1, fig.width=4}
n.size <- map.cols/4

n.cluster <- 2

land.pattern <- matrix(c(1,2,2,1), ncol = 2, nrow = 2)

landclass <- kronecker(
  kronecker(matrix(1,ncol=n.cluster,nrow=n.cluster), land.pattern), 
  matrix(1, nrow=n.size, ncol=n.size))

# ggplot(data.frame(coords, landclass), aes(x = x, y = y, fill = landclass)) +
#   geom_tile(col = "grey50") + theme(legend.position = "none") + labs(title = "landclasses")

```

```{r sim_land_class, fig.asp = 1, fig.width=3}
## 8x8 blocks w/spatial pattern for land cover classes
half.pattern <- matrix(c(rep(c(1,0), each = 4, times = 4), rep(c(0, 1), each = 4, times = 4)), ncol = 8, byrow = TRUE)

quarter.pattern <- matrix(rep(c(rep(c(1,0), each = 2, times = 4), rep(c(0,1), each = 2, times = 4)), times = 2), ncol = 8, byrow = TRUE)

eighth.pattern <-  matrix(rep(c(rep(c(1,0), times = 4), rep(c(0,1), times = 4)), times = 4), ncol = 8, byrow = TRUE)

M.expand <- matrix(1, map.cols/8, map.cols/8)
                   
half.land = kronecker(half.pattern, M.expand)
# image(half.land)
quarter.land = kronecker(quarter.pattern, M.expand)
# image(quarter.land)
eighth.land = kronecker(eighth.pattern, M.expand)
# image(eighth.land)

```

```{r}
Dr <- t(Sigma)
```


```{r sim_one_params}
# x(t) = B1*u1 + B2*u2 + R1*r1 + (p1*u1 + p2*u2 + R2*r2)*t + (g1 + R3*r3)*w1(t) + e(t)

npix = 256
n.time = 6
AR = 0.2
MA = 0.1
burn.in = 1000
B1 = 0.5 # effect of land class 1 over class 0
R1 = 0.001
p0 = 0.3
p1 = 0.2
R2 = 0.001
g1 = 0.2
R3 = 0.001

time = seq(0,1, length.out = n.time)
```

```{r sim_one_errors, }
# spatially correlated errors
del <- t(scale(Dr %*% rnorm(n = npix)))
# ARMA errors (ntime columns)
vareps <- t(rbind(apply(X = del, MARGIN = 2,
                  function(e){
                    arima.sim(n = n.time,
                              model = list(ar = AR, ma = MA, innov = e))
                  })))
# dim(vareps)
```

```{r sim_x}
# B1*u1 + B2*u2 + R1*r1 + (p0 + p1*u1 + p2*u2 + R2*r2)*t + (g1 + R3*r3)*w1(t) + e(t)

landcover = half.land

# Intercepts
land.intercept = if(B1 == 0){rep(0, npix)}else{B1*as.vector(landcover)}
rand.intercept = if(R1 == 0){rep(0, npix)}else{(R1*rnorm(npix))}
int.effect = land.intercept + rand.intercept
int.mat = matrix(rep(int.effect, times = n.time), ncol = n.time)

# Cime effects
land.time = if(p1 == 0){rep(0, npix)}else{p1*as.vector(landcover)}
rand.time = if(R2 == 0){rep(0, npix)}else{(R2*rnorm(npix))}
time.effect = (p0 + land.time + rand.time) %*% rbind(time)

# Climate effects
w.del = t(scale(Dr %*% rnorm(npix)))
w = t(rbind(apply(X = w.del, MARGIN = 2,
                  function(e){
                    arima.sim(n = n.time,
                              model = list(ar = AR, ma = MA, innov = e))
                  })))
clim.par = g1 + if(R3 == 0){rep(0, npix)}else{R3*rnorm(npix)}
clim.mat = matrix(rep(clim.par, times = n.time), ncol = n.time)

clim.effect = clim.mat*w

X = int.mat + time.effect + clim.effect + vareps
```


```{r, fig.asp = 1, fig.width=8, eval = FALSE}
W.data = data.frame(coords, land = as.vector(landcover), W = w) %>% 
  reshape2::melt(id.vars = c("x", "y", "land")) %>% 
  mutate(time = scales::rescale(as.numeric(variable), to = c(0, 1)))

W.data %>%
  ggplot(aes(x = x, y = y, fill = value)) +
  facet_wrap(~variable, ncol = 2) +
  geom_tile() +
  # geom_tile(aes(col = as.factor(land)), alpha = 0.1, fill = NA) +
  scale_fill_viridis_c(option = "magma") +
  # scale_fill_gradient2(low = "brown", high = "forestgreen", mid = "white") +
  labs(col = "land", x = "x coord", y = "y coord", title = "Simulated W") +
  scale_color_manual(values = c("black", "grey80"))

W.data %>% 
  ggplot(aes(x = variable, y = value, col = as.factor(land))) + 
  geom_point() + 
  geom_smooth(method = "lm", formula = y ~ x, aes(group = land), alpha = .2)
```

```{r, fig.asp = 1, fig.width=8}
X.data = data.frame(coords, X, land = as.vector(landcover)) %>% 
  reshape2::melt(id.vars = c("x", "y", "land")) %>% 
  mutate(time = scales::rescale(as.numeric(variable), to = c(0, 1)))

X.data %>%   
  ggplot(aes(x = x, y = y, fill = value, col = as.factor(land))) + 
  facet_wrap(~variable, ncol = 2) + 
  geom_tile() + 
  # geom_tile(aes(col = as.factor(land)), alpha = 0.1, fill = NA) +
  scale_fill_viridis_c(option = "magma") +
  # scale_fill_gradient2(low = "brown", high = "forestgreen", mid = "white") + 
  labs(col = "land", x = "x coord", y = "y coord", title = "Simulated data") + 
  scale_color_manual(values = c("black", "grey80"))
```



```{r, fig.width = 4.5, fig.asp = .8}
tmp <- data.frame(time = time, mean = colMeans(X), sd = apply(X, 2, FUN = sd))

ggplot(tmp, aes(x = time, y = mean)) + 
  geom_point() + 
  geom_line() + 
  geom_smooth(method = "lm", formula = y ~ x) + 
  labs(y = "mean(X)", title = "Overall time trend")

ggplot(X.data, aes(x = time, y = value, col = as.factor(land))) +
  geom_point(alpha = .2, size = 1, position = position_jitter(width = .05)) + 
  # stat_summary(fun.y = "mean", geom = "point") +
  # stat_summary(fun.y = "mean", geom = "line") +
  geom_smooth(method = "lm", formula = y ~ x, aes(group = land), alpha = .2) +
  labs(y = "X", col = "land", title = "Land:time interaction")

ggplot(tmp, aes(x = time, y = sd)) + 
  geom_point() + 
  geom_line() +
  geom_smooth(method = "lm", formula = y ~ x) + 
  labs(y = "sd(X)", title = "Trends in variation")

# ggplot(X.data, aes(x = time, y = value)) +
#   stat_summary(fun.y = "mean", geom = c("line")) +
#   stat_summary(fun.y = "mean", geom = c("point")) + 
#   geom_smooth(method = "lm") + 
#   labs(x = time, y = "mean(X)")

ggplot(data.frame(X = as.vector(X), W = as.vector(w), 
                  land = as.vector(landcover)), 
       aes(x = W, y = X, col = as.factor(land))) +
  geom_point(alpha = .2, size = 1) +
  geom_smooth(method = "lm", formula = y ~ x) +
  labs(col = "land", title = "Effect of climate variable")
```

```{r map_errors, eval = FALSE}
if(npix <= 1000){
  ## Check that errors obtained from mvtnorm behave similarly
  del2 <- mvtnorm::rmvnorm(n = n.time, sigma = Sigma, method = "chol")
  vareps2 <- t(rbind(apply(X = del2, MARGIN = 2,
                           function(e){
                             arima.sim(n = n.time,
                                       model = list(ar = AR, ma = MA, 
                                                    innov = e))
                           })))
  
  library(ggplot2)
  # smooth gradient
  ggplot(data.frame(coords, del), aes(x = x, y = y, fill = del)) +
    geom_tile() +
    scale_fill_viridis_c(option = "magma", name = "delta")
  # less smooth, but obvious spatial correlation
  ggplot(data.frame(coords, del), aes(x = x, y = y, fill = del2)) +
    geom_tile() +
    scale_fill_viridis_c(option = "magma", name = "delta2")
  # slightly checkered
  ggplot(data.frame(coords, vareps[,1]), aes(x = x, y = y, 
                                             fill = vareps[,1])) +
    geom_tile() +
    scale_fill_viridis_c(option = "magma", name = "varepsilon")
  # even more checkered
  ggplot(data.frame(coords, vareps2[,1]), aes(x = x, y = y, 
                                              fill = vareps2[,1])) +
    geom_tile() +
    scale_fill_viridis_c(option = "magma", name = "varepsilon2")
  
  ## Conclusion, the mvtnorm method produces LESS spatially correlated data
  ## than t(Dr %*% rnorm())
}
```

# Simulations

The actual simulations were run with the script "self-contained-batch-sim_script/run-PARTS-sim.R"

```{r}
# B1*u1 + B2*u2 + R1*r1 + (p0 + p1*u1 + p2*u2 + R2*r2)*t + (g1 + R3*r3)*w1(t) + e(t)
# X(t) = lambda*land + S1*r1 + (alpha + beta*land + S2*r2 )*t + (gamma + S3*r3)*temp(t) e(t)
# 
```

1) Spatial Only

  * T = 1
  
  * varying map sizes: e.g., 10K, 20K, 40K, 80K
  
  * fixed extent of **u**
  
  * test that beta1 = 0 and beta2 = beta1
  
  * plot estimates of beta1 with standard errors
  
  * **p** = **R2** = **g** = **R3** = 0
  
```{r}
source("scripts/simulation-functions_morrowcj.R")
sims = 10 # start small to test
```

With this first simulation we ask "Does a map size affect coefficient estimates?"
  
```{r sim_1}
if(TRUE){
  sim1_time <- system.time(expr = {
    sim1 = sim_PARTs(map.width = c(104, 144, 200, 280), # varying map size
                     ntime = 1, # spatial only
                     land.grid = 2, # fixed land class extent.
                     lambda = .2, S1 = 0, # constant effect of land class
                     alpha = 0, beta = 0.2, S2 = 0, # effect of time in beta only
                     gamma = 0, S3 = 0, # no effect of climate
                     r = .05, # constant spatial range of ~ <5%
                     nug = 0.1, # constant nugget
                     AR = .2, MA = 0, # constant AR and MA parameters
                     fixed_temp = TRUE, fixed_r2 = TRUE, fixed_r3 = TRUE, # fix to avoid recalculating
                     par.sims = sims, # number of simulations per map
                     npart = NA, partsize = 2000, # partitions are 2000px for each map size
                     GLS.resp = "time", # estimate effects on time in GLS
                     ncores = 10, verbose = TRUE, debug = FALSE
    )
  })
  sim1_time
  save(sim1, sim1_time, file = "data/sim1.RData")
} else{
  load("data/sim1.RData")
}
```

I need to fix the r estimation.

```{r, fig.asp = 1}
library(ggplot2)
L = sim1
df = L$sim.table %>% cbind(lapply(L$GLS, function(x){x$overall_stats$coefmean}) %>% bind_rows())
df.melt <- df %>%
  # mutate(time.eff = `(Intercept)` + land1)
  reshape2::melt(measure.vars = grep(pattern = "(\\(Intercept\\))|(^land1)",
                                     x = names(df), value = TRUE),
                 variable.name = "coefficient", value.name = "estimate")

df.melt %>% 
  ggplot(aes(x = factor(map.width), y = estimate, col = coefficient)) +
  theme_classic() +
  geom_hline(yintercept = c(0, unique(df.melt$beta)), lty = 2, col = "grey50") +
  stat_summary(size = .25) + stat_summary(geom = "line", aes(group = coefficient)) +
  labs(x = "Map Width", y = "Estimate")
```
  
  
2) Variable land patch sizes

  * T = 30
  
  * map size = 10K
  
  * variable extent of **u**
  
  * test that p1 = 0 and p1 = p2
  
  * **g** = **R3** = 0
  
3) Effect of spatio-temporal predictor 

  * T = 30
  
  * map size = 10K
  
  * fixed extent of **u**
  
  * test that g1 = 0
  
  * **p** = **R2** = 0
  
  * variable **R2** if spatial mean of **r2** is 0.
  
4) Effect on Type I errors (p1 = 0 and p1 = p2)

  * variable **w**, all else constant
  
  * variable **r2**, all else constant
  
  * variable **r1**, all else constant
  
  * variable range parameter in **e(t)**, all else constant
  
  * spatially variable **rho** (2D sin wave), all else equal
