---
title: "PARTS Simulations"
author: "Clay Morrow"
date: "3/10/2021"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE, cache.extra = rand_seed)
```

```{r, message=FALSE, warning=FALSE, error=FALSE}
library(mvtnorm)
library(dplyr)
library(ggplot2)
```


<!-- ## Original Models -->

<!-- Simulation model:  -->
<!-- $x_i(t) = (\Sigma_{k} \nu_{k, i} u_{k, i}) + \omega_i w_{i}(t) + (\Sigma_{k} \zeta_{k,i} u_{k,i}) t + \varepsilon_{i}(t)$ -->

<!-- with `ARMA(2, 1)` error structure:  -->
<!-- $\varepsilon_{i}(t) = \beta_{1,i} \varepsilon_{i}(t - 1) + \beta_{2,i}\varepsilon_{i}(t - 2) + \delta_i(t) + \alpha_{i} \delta_{i}(t - 1)$ -->

<!-- where:  -->

<!-- * $x_t(t)$: spatio-temporal response (e.g. NDVI) at location $i$ at time $t$ -->

<!-- * $w_i(t)$: spatio-temporal predictor (e.g temperature) -->

<!-- * $u_{k,i}$: $k$ model matrix of temporally-fixed spatial variable (e.g.  -->
<!-- land-cover); $k = 0, ..., K$. When $k = 0$, $u$ is a vector of ones  -->
<!-- (intercept only) -->

<!-- * $(\Sigma_{k} \zeta_{k, i} u_{k, i})$: time trend term    -->

<!-- * $(\Sigma_{k} \nu_{k, i} u_{k, i})$: Intercept term -->

<!-- * $\delta_i(t)$: random variation driver: multivariatiate normal -->

<!--   - $cor{\delta_i(t), \delta_j(t)} = f_\delta(d_{i,j}; \psi_i, r_{i,j}, a_{i,j})$ -->

<!--     + $f_\delta()$ is some function that depends on distance between points  -->
<!--     $i$ and $j$.  -->

<!-- ```{r stackex_arima_sim, eval = FALSE, echo = FALSE} -->
<!-- # I should consider generating $\varepsilon_{i}(t)$ with the `arima.sim()`  -->
<!-- # function. This would allow easy generalization to any ARMA(p, q) process.  -->
<!-- # I'd need to figure out how to make time series correlated. The `arima.sim()`  -->
<!-- # allows use of a function to generate the error term. An answer [here](https://stats.stackexchange.com/questions/136462/how-do-you-generate-correlated-arma1-1-models)  -->
<!-- # shows how to do that: -->


<!-- ## Code from StackExchange -->
<!-- ## Here, X and Y are correlated ARMA(1, 1) -->
<!-- ## x(t) = phi*x(t-1) + e(t) + theta*e(t-1) -->
<!-- ## y(t) = alpha*y(t-1) + eta(t) + beta*eta(t-1) -->
<!-- ## cov(e(i), eta(i)) = sigma; cov(e(i), eta(j)) = 0 -->


<!-- # require("mvtnorm") -->
<!-- # parameters of the data generating process -->
<!-- desired.rho <- 0.2 -->
<!-- Sigma <- diag(c(2,3))     -->
<!-- phi <- 0.7 -->
<!-- theta <- 0.3 -->
<!-- alpha <- -0.4 -->
<!-- beta <- 0.2 -->
<!-- # implied value of the covariance between the disturbance terms -->
<!-- sdx <- sqrt(Sigma[1,1] * (1 + theta^2 + 2*phi*theta) / (1 - phi^2)) -->
<!-- sdy <- sqrt(Sigma[2,2] * (1 + beta^2 + 2*alpha*beta) / (1 - alpha^2)) -->
<!-- scov <- desired.rho * sdx * sdy * (1 - phi*alpha) / (phi*beta + 1 + alpha*theta + beta*theta) -->
<!-- scov -->
<!-- # [1] 1.022579 -->
<!-- Sigma[1,2] <- Sigma[2,1] <- scov -->
<!-- n <- 200 -->
<!-- niter <- 10000 -->
<!-- rhos <- rep(NA, niter) -->
<!-- set.seed(123) -->
<!-- for (i in seq_len(niter)) { -->
<!--   eps <- mvtnorm::rmvnorm(n = n, mean = rep(0, 2), sigma = Sigma) -->
<!--   x1 <- arima.sim(n = n, model = list(ar = phi, ma = theta), innov = eps[,1]) -->
<!--   x2 <- arima.sim(n = n, model = list(ar = alpha, ma = beta), innov = eps[,2])   -->
<!--   rhos[i] <- cor(x1, x2) -->
<!-- } -->
<!-- summary(rhos) -->

<!-- ## Notes:  -->


<!-- # As a varcovar matrix, the diagonal of Sigma is the variance for each variable. -->
<!-- ``` -->


<!-- ### Notes for Tony -->

<!-- * Why is $u$ a vector of ones when $k = 0$? -->
<!-- * What is $\Sigma_k$ in the intercept and time-trend parameters? -->



# Reworked models 

The overall response will be simulated with the following process:

$$
x_{i}(t) = \beta_{k, i}u_{k, i} + (\phi_{m, i}u_{m,i})t + 
\gamma_{k, i} w_{k, i}(t) + \varepsilon_{i}(t) \tag{1}
$$

Where: 

* $x_{i}(t)$ is the variable of interest (e.g. NDVI) in pixel $i$;

* $u_{k,i}$ is a model matrix of $k = 1, ..., K$ spatially variable predictors 
(e.g. landclass);

* $w_{m, i}(t)$ are $m = 1, ..., M$ spatio-temporally variable predictors 
(e.g. temperature);

* $\beta_{k, i}$, $\phi_{m, i}$, and $\gamma_{m, i}$ are the effects of the 
predictors. Each of these can be fixed OR randomly distributed in space 
$\{\beta_{k}, \phi_{m},  \gamma_{m} \} \sim N(\mu, \Sigma_\text{Matern})$ 
(see below); and

* $\varepsilon_{i}(t)$ is the spatio-temporally correlated error term.

$\varepsilon_{i}(t)$ (and $w_{m,i}$) is generated with an ARMA process:

$$\varepsilon_{i}(t) = \delta_{i}(t) + 
\sum_{j = 1}^{p}{\rho_{j}\varepsilon_{i}(t)} + 
\sum_{j = 1}^{q}{\theta_{j}\varepsilon_{i}(t)} \tag{2}$$

where: 

* $\sum_{j = 1}^{p}{\rho_{j}\varepsilon_{i}(t)}$ is the AR(p) process;

* $\sum_{j = 1}^{q}{\theta_{j}\varepsilon_{i}(t)$ is the MA(q) process; and

* $\delta_{i}(t) \sim N(0, \Sigma_{\delta_{i}})$ are spatially correlated 
errors;

Spatial correlation in the ARMA terms is generated with an exponential power 
function: 

$$\Sigma_{\delta_{i}} = I\eta_i + (\sigma^{2} - \mu_{\eta_i})exp{((-d_{i, j}/r)^a)} \tag{3}$$

where: 

* $I\eta_{i}$ is an Identity matrix multiplied by the nugget $\eta_{i}$, 
which may be fixed OR variable in space: 
$\eta_{i} \sim N(0, \Sigma_{\text{Matern}})$; 

* $\sigma^2$ is the sill (variance) of $\delta_i$ (I'll stick with unit variance);

* $d_{i, j}$ is the distance between pixels $i$ and $j$;

* $r$ is the range of the spatial auto-correlation (in pixels); and

* $a$ is the shape parameter for the auto-correlation function

Other spatial correlations $\Sigma_{\text{Matern}} = R(h)$ will be generated the 
Matern Cross-covariance function: 

$$R(d_{i, j}; v, c) = 2^{1-v} \Gamma(v)^{-1}(c||d_{i, j}||)^{v} K_{v}(c||d_{i, j}||) \tag{4}$$
where :

* $c > 0$ is the scale parameter (similar to range $r$)

* $v > 0$ is the smoothness parameter. When $v = \frac{1}{2}$, 
$R(h)$ is equivalent to $exp(d_{i, j} / c)$ and $\lim_{v \to \infty} R(h) = exp(-d/2c)^2$

* $\Gamma$ is the gamma function, and 

* $K$ is a modified Bessel function of the second kind

**Note:** In practice, this function is very useful for generating smooth spatial 
correlations and waves (see $\S$ "Spatially variable parameter" below). 
But may need some help understanding the math and what it's actually 
doing! It may be worth while to also generate $\delta_{i}$ with Matern correlations?

# Simulation Plans

## Data Generation

In order to vigorously test the robustness of the PARTS method, we will simulate
1000 sets of predictor variables $u_{k,i}$ and $w_{m,j}$ and varying the effects
that each of those sets have on $x_i(t)$. Therfore, $x$ will be simulated
$1000$ for each unique set of effect parameter combinations. The map for all 
simulations will be the same size: $20\times20$ pixels (or some other size, 
dependant on speed). $d_{i, j}$ will contain the Euclidean distance between 
points and will be constant throughout the simulations. 

$u$ will represent land cover class designations, with 4 levels ($m = 4$), coded as dummy 
variables (i.e. $\{0, 1\}$). It will be generating by discretizing (by quartiles)
a continuous variable produced with the Matern covariance function. $c$ and $v$
will be selected to generate a desired spatial pattern and will be constant 
throughout the simulations. $u$ will have no nugget and unit variance 
(i.e. $\eta = 0$, $\sigma^2 = 1$).

$w$ will be a single ($k = 1$) spatio-temporal variable representing temperature.
It will be generated as an ARMA(1, 1) model with Matern covariance structure. 
$c$ and $v$ will be selected to generate a desired spatial pattern 
(different from $u$) and will be constant throughout the simulations. $\rho$ and 
$\theta$ will also be predetermined and constant. $w$ will have unit variance.
$w$ will have no nugget and unit variance (i.e. $\eta = 0$, $\sigma^2 = 1$).

$x$ will then be generated, for 30 time points ($t = 1, ..., 30$), by varying 
the following parameters on each of the 1000 data sets. This will allow us to
compare the effect of the parameter independent of the specific predictor values.

Effects parameters $\beta$, $\phi$, and $\gamma$, will take on one of 3 levels: 2 fixed values 
(i.e. $\{0, 0.5\}$) and one spatially distributed random variable
(i.e. $\sim N(0.5, \Sigma_{\text{Matern}})$). The AR parameter $\rho$ will also have 2 levels (e.g. $\{\rho_1 = .6, \rho_2 = .3\}$ and $\{\rho_1 = .8, \rho_2 = .1\}$), the MA parameter $\theta$ 
will have 2 levels (e.g $\{.1, .5\}$), the spatial nugget $\eta$ will
have 2 levels (i.e. $\{0, 0.5\}$), and the spatial correlation range parameter 
$r$ will have 2 levels (i.e. $\{0.2, 0.5\}$). The spatial correlation shape 
parameter $a = 2$ will remain constant througout the simulations.

With this specification there are 432 unique parameter combinations which would
yield 432,000 simulations of $x$. It may be best to also hold $\beta$ constant 
or at least restrict it to 2 fixed values, since we aren't actually interested 
in our estimate of an the intercept parameter. Doing so would reduce the total
simulations to 72,000. It may also be acceptable to hold $\theta$ constant, 
further reducing the total to 36,000 simulations. But, we may also decide to 
add a level of $\eta$ that is spatially variable.

## Model Fitting

For each simulation of $x$, we will apply the PARTS method. We will use AR_REML
to fit a time series to each pixel and then will fit a GLS to slope parameters 
for the effect of time. We will then compare the overall estimate of the effect 
of time to the true values $(\phi_{m, i})$. If $\phi_m$ was spatially variable, 
we will compare to the mean $\phi_m$. We will also estimate type I error rates 
for the case when $\bar{\beta_m} = \bar{\phi_m} = \bar{\gamma_k} = 0$.

# Example Code 

This is some code I ran to test out the spatial correlation functions

First we'll set the initial parameters:

```{r start_params}
#---
## Initial Parameters
#---

set.seed(3654)

# Sim pars ----
burn.in = 0
sim.iters = 1

# Map pars ----
map.cols = 20
map.rows = 10
npix = map.rows * map.cols # should be a square?

# ARMA pars ----
rhos = c(0.6, 0.2)# AR parameter(s): must sum to less than 1
thetas = c(.4) # MA parameter: must be less than 1

# Temporal pars ----
ntime = 30 # number of time points

# varcov pars ----
cormod = "stable" # 'stable' correlation model: varcov = nug+sill*exp(-(d/scale)^shape)
dist.meth = "Eucl" # Euclidean distance
sill = 1 # sill (variance)
r = .25 # range of spatial autocor, relative to width of map (scale parameter)
a = 1 # shape of spatial autocor function (shape parameter)
nug = .2 # fixed nugget

# spatial nugget pars ----
nug.cormod = "matern"
nug.distmeth = "Eucl"
nug.r = .4
nug.s = 20
constrain.nug = c(0, 1) # constrain the spatial nugget to these values

# effect parameters ----
beta = .2 # spatial predictors
phi = .2 # spatio-temporal predictors
gamma = .5 # time trends

# Additinal parameter calculations ----

## Assign coordinates to pixels
coords = as.matrix(expand.grid(x = as.double(1:map.cols), 
                               y = as.double(1:map.rows)))

## spatial correlation
psill = sill - nug # partial sill: if nug is spatially variable, use mean(nug)
range = r * map.cols # range in pixels

## spatially variable parameters
nug.range = nug.r * map.cols
nug.smooth = nug.s


# Sanity checks ----
stopifnot(sum(rhos) < 1)
stopifnot(nrow(coords) == npix)

# parameter table ----

fixed.pars = expand.grid(range = range, a = a, nug = nug, fixed.pars = TRUE)


corpars = list(scale = range, power = a, # parameter list for Covmatrix()
               sill = psill, nugget = nug, mean = 0)
nugpars = list(scale = nug.range, smooth = nug.smooth, 
               mean = 0, sill = 1, nugget = 0)
```

Then:

1) Calculate the spatial covariance matrix $\Sigma_{\delta_{i}}$ with the
function `Covmatrix` from package `CompRandFld`.

```{r covmat, eval = FALSE}
start.time = Sys.time()

# Calculate varcov matrix (Sigma) ----
## Covmatrix() - easier to modify specs
covmat = CompRandFld::Covmatrix(coordx = coords, 
                                 distance = dist.meth,
                                 corrmodel = cormod,
                                 param = corpars)$covmatrix

covmat.tap = CompRandFld::Covmatrix(coordx = coords, 
                                 distance = dist.meth,
                                 corrmodel = cormod, 
                                 type = "Tapering", taper = "Wendland1", 
                                 param = corpars)$covmatrix

# invcholV = remotePARTS::invert_chol(covmat)

# ## Tony's
# d = as.matrix(stats::dist(coords, method = "euc"))
# d.scale = (d - min(d)) / (max(d) - min(d)) # rescale to (0, 1)
# covmat2 = diag(rep(nug, npix)) + (1 - nug) * exp(-(d/r)^a)
# invcholV2 = remotePARTS::invert_chol(covmat2)
# ## Same answer?
# all((covmat2 - covmat) == 0) # TRUE! - also no speed change
```

```{r test tapering functions}
## Wendland tapering function
Wendl <- function(h, theta = 0.2){
  w = ifelse(test = abs(h) < theta, 
             yes = 0,
             no = (1 - (abs(h)/theta))^4 * (1 + (4*(abs(h)/theta))))
  
  return(w)
}

## Spherical tapering function
taper.sphr <- function(d, theta, type = "var") {
  if(type == "dist"){
    ts = ifelse(d > theta, 
           0, 
           ((1 - d/theta)^2) * (1 + d/(2 * theta))
    )
  }
  if(type == "var"){
    ts = ifelse(d < theta,
           0,
           ((1 - d/theta)^2) * (1 + d/(2 * theta)))
  }
  return(ts)
}
```

```{r, }
d = scales::rescale(as.matrix(dist(as.matrix(coords))), to = c(0, 1))
prop.zero = .6
wend.range = quantile(covmat, probs = prop.zero) # 25% of cells should be zero

# 
# image(covmat)
# 
# image(Wendl(covmat, theta = wend.range))
# image(taper.sphr(covmat, .5, type = "var"))
# image(taper.sphr(d, .5*map.cols))
```

```{r, eval = FALSE}
## What does theta need to be to have relatively many zeros?
zeros = Vectorize(function(r){sum(taper.sphr(d, r, type = "dist") == 0)/length(d)})
curve(expr = zeros(x), from = 0, to = 1)
x.tmp = seq(0, 1, length.out = 101)
y.tmp = zeros(x.tmp)

# fm <- lm(y.tmp ~ x.tmp + I(x.tmp^2) + I(x.tmp^3) + I(x.tmp^4))
# coefs = fm$coefficients
# 
# plot(x = x.tmp, y = y.tmp);lines(x = x.tmp, y = c(1, predict(fm)), col = "red")

# 80% zeros 
max(x.tmp[y.tmp >= .6], na.rm = TRUE) #theta <= 0.33
max(x.tmp[y.tmp >= .7], na.rm = TRUE) #theta <= 0.26
max(x.tmp[y.tmp >= .8], na.rm = TRUE) #theta <= 0.21
```

```{r}
covmat.sphr = taper.sphr(covmat, wend.range)
covmat.wendl = covmat.sphr = Wendl(covmat, wend.range)
covmat.sphr.d = diag(rep(nug, npix)) + (1 - nug) *taper.sphr(d, theta = r, type = "dist")
sum(covmat.sphr.d == 0)/length(d)

# system.time(t(chol(covmat.sphr.d)))
# 
# covmat.sprs = as.spam(covmat.sphr.d)
# system.time(t(chol(covmat.sprs))) #took 64.5% as long
# 
# library(Matrix)
# covmat.M = as(covmat.sphr.d, "dgTMatrix")
# system.time(t(Matrix::chol(covmat.M))) #took 92.2% as long
```

```{r fig.asp = .8, fig.width = 3.5}
rv = rnorm(npix, sd = 0.5)
test.dat = data.frame(coords, 
                      z.og = t(chol(covmat)) %*% rv,
                      z.og.taper = t(chol(covmat.tap)) %*% rv,
                      z.sphere = t(chol(covmat.sphr)) %*% rv,
                      z.wendl = t(chol(covmat.wendl)) %*% rv,
                      z.sphr.d = t(chol(covmat.sphr.d)) %*% rv
                      )

test.dat %>% 
  ggplot(aes(x = x, y = y, fill = z.og)) + 
  geom_tile() + 
  # theme(legend.position = "none") +
  scale_fill_viridis_c(option = "magma") + 
  labs(title = "Original exp-pow Covariance")

# test.dat %>%
#   ggplot(aes(x = x, y = y, fill = z.og.taper)) +
#   geom_tile() +
#   # theme(legend.position = "none") +
#   scale_fill_viridis_c(option = "magma") +
#   labs(title = "Original Wendland1 Taper")
# 
# test.dat %>%
#   ggplot(aes(x = x, y = y, fill = z.wendl)) +
#   geom_tile() +
#   # theme(legend.position = "none") +
#   scale_fill_viridis_c(option = "magma") +
#   labs(title = "Wendland Tapered Matern")
# 
# test.dat %>%
#   ggplot(aes(x = x, y = y, fill = z.sphere)) +
#   geom_tile() +
#   # theme(legend.position = "none") +
#   scale_fill_viridis_c(option = "magma") +
#   labs(title = "Spherically Tapered Matern")

test.dat %>% 
  ggplot(aes(x = x, y = y, fill = z.sphr.d)) + 
  geom_tile() + 
  # theme(legend.position = "none") +
  scale_fill_viridis_c(option = "magma") + 
  labs(title = "Spherical Taper")
```

```{r, fig.asp = .8, fig.width = 3.5}
# Dr <- as(t(chol(exp(-Dist/r))), "dgTMatrix")
# e <- Dr %*% rnorm(n, sd = s)

Dr = t(chol(covmat.sphr.d))
e = scale(Dr %*% rnorm(npix, sd = .5))

data.frame(coords,e = e) %>% 
  ggplot(aes(x = x, y = y, fill = e)) + 
  geom_tile() +
  scale_fill_viridis_c(option = "magma") 
```


```{r}
prop.zero = .8
wend.range = quantile(covmat, probs = prop.zero) # 25% of cells should be zero

tap.theta = quantile(d, 1-prop.zero)# 25% of cells should be zero

sum(taper.sphr(d, tap.theta) == 0)/length(d)

image(Wendl(covmat, wend.range))
image(taper.sphr(covmat, 0.25, type = "var"))
image((exp(-d/10))) # non-tapered 
# image(taper.sphr(d, tap.theta))


```

2) generate $\delta_i$ with `rmvmnorm()`.

```{r space_err, eval = FALSE}
# Generate spatially correlated errors (delta) ----
del = mvtnorm::rmvnorm(n = ntime, sigma = covmat, method = "chol")
```

3) generate $\varepsilon_{i}(t)$.

```{r spacetime_err}
# Generate ARMA errors (epsilon) ----
## apply arima.sim() to each column of del
eps = t(apply(X = del, MARGIN = 2, 
             function(e){
               arima.sim(n = ntime, 
                         model = list(ar = rhos, ma = thetas), 
                         innov = e)
             }))

end.time = Sys.time()
```

```{r}
end.time - start.time # 3.42 minutes
```

Let's visualize the correlations among $\varepsilon_{i}(t)$ in space and time:

```{r viz_err, fig.asp = 1, fig.width= 3.5, echo = FALSE}
# Visualize temporal and spatial correlation ----
cor(eps) %>% reshape2::melt() %>% 
  ggplot(aes(x = Var1, y = Var2, fill = value)) + 
  geom_tile() +
  scale_fill_viridis_c(option = "magma") +
  theme(legend.pos = "none") +
  labs(x = "time i", y = "time j", title = "Temporal correlation")


cor(t(eps)) %>% reshape2::melt() %>% 
  ggplot(aes(x = Var1, y = Var2, fill = value)) + 
  geom_tile() +
  scale_fill_viridis_c(option = "magma") +
  theme(legend.pos = "none") +
  labs(x = "pixel i", y = "pixel j", title = "Spatial correlation")
```

4) Add the errors to the rest of the sim model

```{r, eval = FALSE}
# beta, phi, gamma: user defined effects parameters (possibly spatially variable)
# u: spatially generated predictors (via mattern function)
# m: spatio-temporally generated predictors (mattern & ARMA functions)

x = beta*u +  phi*u*time  + gamma*w + eps 

#Note: proper matrix multiplication will be used
```

# Spatially variability: Matern Covariance

To demonstrate how we can simulate a waved spatial pattern in our predictors 
and effects parameters, I'll now generate a spatially variable nugget using 
Matern cross-covariance. When $v$ and $c$ are both large, you get a wave
across the map:

```{r, fig.asp = .8, fig.width = 3.5, echo = FALSE}
## spatially distributed nugget?
# ?CompRandFld::Covmatrix(coords, distance = dist.meth, cormod = "wave")
# ?CompRandFld::CorrelationParam(corrmodel = "wave")

param.vcov = CompRandFld::Covmatrix(coords, 
                       distance = dist.meth, 
                       corrmodel = "matern",
                       param = nugpars)$covmatrix

## Tapered version: Only 1s and 0s
# param.vcov = CompRandFld::Covmatrix(coords, 
#                        distance = dist.meth, 
#                        corrmodel = "matern", type = "Tapering", taper = "Wedland2",
#                        param = nugpars)$covmatrix

param.corpars = list(scale = range, power = a, sill = psill, nugget = 0, mean = 0)
covmat = CompRandFld::Covmatrix(coordx = coords, 
                                 distance = dist.meth,
                                 corrmodel = cormod,
                                 param = param.corpars)$covmatrix

spat.param <- scales::rescale(mvtnorm::rmvnorm(n = 1, sigma = param.vcov), to = constrain.nug)

ggplot(data = data.frame(coords, param = as.vector(spat.param)), 
       aes(x = x, y = y, fill = param)) +
  theme_classic() + 
  theme(legend.position = "none") +
  geom_tile() + 
  scale_fill_viridis_c(option = "magma") +
  labs(x = "x coord", y = "y coord", title = "spatially variable patameter")

## If this parameter is the nugget, then you can use it to simulate ARMA errors:
covmat = diag(as.vector(spat.param)) + (1-mean(as.vector(spat.param)))*covmat
del = mvtnorm::rmvnorm(n = ntime, sigma = covmat, method = "chol")
eps = t(apply(X = del, MARGIN = 2, 
             function(e){
               arima.sim(n = ntime, 
                         model = list(ar = rhos, ma = thetas), 
                         innov = e)
             }))
```

The Matern function is ideal for simulating spatial variables and spatial parameters
because the resultant patterns can be **substantially** changed by varying $v$ and $c$. 
To demonstrate this, I'll vary $c$ and $r$. $c = {1/2, 1, 10, 20}$ and
$v = {1/2, 1, 2, 10}$

```{r, fig.asp = 1, fig.width = 6, echo = FALSE}
c.tmp = c(.5, 1, 2, 10)

v.tmp = c(1/2, 1, 5, 10)

spatcor.tab = data.frame(NULL)

for (v in v.tmp) for (c in c.tmp){
  
  nugpars = list(scale = c, smooth = v, 
               mean = 0, sill = 1, nugget = 0)
  
  nug.vcov = CompRandFld::Covmatrix(coords, 
                                    distance = dist.meth, 
                                    corrmodel = "matern", 
                                    param = nugpars)$covmatrix
  # image(nug.vcov)
  spat.nug <- scales::rescale(mvtnorm::rmvnorm(n = 1, sigma = nug.vcov), to = constrain.nug)
  tmp = data.frame(coords, nug = as.vector(spat.nug), v = v, c = c)
  spatcor.tab = rbind(spatcor.tab, tmp)
  
  # nug.corpars = list(scale = range, power = a, sill = psill, nugget = 0, mean = 0)
  # covmat = CompRandFld::Covmatrix(coordx = coords, 
  #                                 distance = dist.meth,
  #                                 corrmodel = cormod,
  #                                 param = nug.corpars)$covmatrix
  # 
  # covmat = diag(as.vector(spat.nug)) + (1-mean(as.vector(spat.nug)))*covmat
  # del = mvtnorm::rmvnorm(n = ntime, sigma = covmat, method = "chol")
  # eps = t(apply(X = del, MARGIN = 2, 
  #               function(e){
  #                 arima.sim(n = ntime, 
  #                           model = list(ar = rhos, ma = thetas), 
  #                           innov = e)
  #               }))
  
}

ggplot(data = spatcor.tab, aes(x = x, y = y, fill = nug)) +
  theme_classic() + 
  theme(legend.position = "none",
        panel.spacing = unit(0, "lines")) +
  geom_tile() + 
  scale_fill_viridis_c(option = "magma") +
  facet_grid(c ~ v, labeller = "label_both") +
  labs(x = "x coord", y = "y coord", title = "spatial patterns (matern covariance)")
```

whereas the stable exponential function has more limited range ($0<a\leq2$). 
Though, the Gaussian version ($a = 2$) can generate interesting and realistic
patterns:

```{r, fig.asp = 1, fig.width = 6}
r.tmp = c(.05, 0.2, .5, 1)*map.cols

a.tmp = c(.5, 1, 1.5, 2)

spatcor.tab = data.frame(NULL)

for (a in a.tmp) for (r in r.tmp){
  
  nugpars = list(scale = r, power = a, 
               mean = 0, sill = 1, nugget = 0)
  
  nug.vcov = CompRandFld::Covmatrix(coords, 
                                    distance = dist.meth, 
                                    corrmodel = "stable", 
                                    param = nugpars)$covmatrix
  # image(nug.vcov)
  spat.nug <- scales::rescale(mvtnorm::rmvnorm(n = 1, sigma = nug.vcov), to = constrain.nug)
  tmp = data.frame(coords, nug = as.vector(spat.nug), r = r/map.cols, a = a)
  spatcor.tab = rbind(spatcor.tab, tmp)
  
  # nug.corpars = list(scale = range, power = a, sill = psill, nugget = 0, mean = 0)
  # covmat = CompRandFld::Covmatrix(coordx = coords, 
  #                                 distance = dist.meth,
  #                                 corrmodel = cormod,
  #                                 param = nug.corpars)$covmatrix
  # 
  # covmat = diag(as.vector(spat.nug)) + (1-mean(as.vector(spat.nug)))*covmat
  # del = mvtnorm::rmvnorm(n = ntime, sigma = covmat, method = "chol")
  # eps = t(apply(X = del, MARGIN = 2, 
  #               function(e){
  #                 arima.sim(n = ntime, 
  #                           model = list(ar = rhos, ma = thetas), 
  #                           innov = e)
  #               }))
  
}

ggplot(data = spatcor.tab, aes(x = x, y = y, fill = nug)) +
  theme_classic() + 
  theme(legend.position = "none",
        panel.spacing = unit(0, "lines")) +
  geom_tile() + 
  scale_fill_viridis_c(option = "magma") +
  facet_grid(r ~ a, labeller = "label_both") +
  labs(x = "x coord", y = "y coord", title = "spatial patterns (stable exp covariance)")
```

The matern covariance structure seems to have been developed to generate more
bioligically realistic spatial patterns. 
